---
title: "Project proposal"
author: "THE Triangle"
output: github_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(readxl)
library(janitor)
library(broom)
```

## 1. Introduction
Microplastics have become more and more well known throughout the world in the past few years. We now recognize that they are an incredibly dangerous bi-product to marine life and many organizations have begun to attack the problem. That is what this project is about. Our project will be looking at a 2023 dataset by Leslie Hart of microplastic samples in prey fish caught in Sarasota Bay. The samples were extracted from muscle tissue and GI tracts of the fish, and sorted by size, type, color, and quantity.

Questions:
Is there a discrepancy in the amount of microplastics in different locations of the fish?
Does the size of the fragment have a relationship to the location? Are larger fragments found in the muscle or in the GI tract?



## 2. Data

```{r import-and-clean-data, message=FALSE}
fish_microplastics <- read_xlsx("../data/Fish_Microplastics_Data_Repository.xlsx", 
                                na = "NA") %>%
  janitor::clean_names() 

fish_microplastics <- fish_microplastics %>%
  mutate(sieve_size = str_extract(sieve_size, pattern = "[:digit:]+")) %>%
  mutate(sieve_size = case_when(sieve_size == "LG" ~ 777,
          TRUE ~ as.numeric(sieve_size))) %>%
  mutate(hot_needle = case_when(hot_needle == "yes" ~ TRUE, hot_needle == "no" ~ FALSE))
```
```{r glimpse}
glimpse(fish_microplastics)
```
## 3. Ethics review


## 4. Data analysis plan
With the data we plan on plotting it in a spatial manner. We want to see where these fish were gathered and if there is any correlation. We also want to see if there is a correlation between different fish and the average count of micro plastics found within it. It would be interesting to see if multiple variables like the distinct IDs and the average count have any correlation.


