---
title: "Project proposal"
author: "THE Triangle"
output: github_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(readxl)
library(janitor)
library(broom)
library(ggridges)
library(leaflet)
library(htmltools)
library(sf)
```

## 1. Introduction
Microplastics have become more and more well known throughout the world in the past few years. We now recognize that they are an incredibly dangerous byproduct to marine life and many organizations have begun to investigate the problem. That is what this project is about. Our project will be looking at a 2023 dataset by Leslie Hart of microplastic samples in prey fish caught in Sarasota Bay. The samples were extracted from muscle tissue and GI tracts of the fish, and sorted by size, type, color, and quantity.

Questions:
Is there a discrepancy in the amount of microplastics in different locations of the fish?
Does the size of the fragment have a relationship to the location? Are larger fragments found in the muscle or in the GI tract?



## 2. Data

```{r import-and-clean-data, message=FALSE}
fish_microplastics <- read_xlsx("../data/Fish_Microplastics_Data_Repository.xlsx", 
                                na = "NA") %>%
  janitor::clean_names() 

fish_microplastics <- fish_microplastics %>%
  mutate(sieve_size = str_extract(sieve_size, pattern = "[:digit:]+")) %>%
  mutate(sieve_size = case_when(sieve_size == "LG" ~ 777,
          TRUE ~ as.numeric(sieve_size))) %>%
  mutate(hot_needle = case_when(hot_needle == "yes" ~ TRUE, hot_needle == "no" ~ FALSE))
```
```{r glimpse}
glimpse(fish_microplastics)
```
## 3. Ethics review
Besides from people choosing to eat the species of fish that were studied, there is very little potential negative impact on people. The spatial data is very limited because the data was only collected at two sampling stations. The data is not characteristic of the entire ecosystem because they only studied 4 species, we would have to do more research into whether they are widely predated or not. This is dependent on ongoing research about micro plastics and whether or not they are heavily accumulated by trophic transfer. A potential positive impact is that if our data shows that fish from a certain area are particularly contaminated with microplastics, it could fuel cleanup efforts in that area. If anything, the impact would be positive because people would be more aware of potentially contaminated species, but any impact on individuals, positive or negative, will most likely marginal.

## 4. Data analysis plan

With the data, we plan on plotting it in a spatial manner using Leaflet. We want to see "in relation to the species type, is there any correlation to the quantity and sample mass of the microplastic particle type?". As well as asking the question of "does the color of the microplastic lend any correlation to the distribution and placement in the sample type?". The combination of these to questions might allow us to understand if there correlation between different fish and the average count of microplastics found within Sarasota Bay. It would be interesting to see if multiple variables like the distinct IDs and the average count have any correlation.


```{r prelim-plot-gi-bar, fig.alt="This bar graph shows the distribution of microplastics within the fish. We can see that the majority of the microplastics can be found in the GI tract."}
fish_microplastics %>%
  group_by(fish_id) %>%
  ggplot(aes(x = fish_id, fill = sample_type)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```
```{r prelim-plot-quant-density, fig.alt="In this ridgeline plot, we can see the different individual fish that were collected and the quantity of microplastics within them. We can see that most of the fish had a similar density of microplastics at the lower end of the scale, with some individuals having spikes of concentration at higher levels."}
fish_microplastics %>%
  group_by(fish_id) %>%
  filter(quantity < 60) %>%
  ggplot(aes(x = quantity, y = fish_id, fill = fish_id)) +
  geom_density_ridges()
```
```{r prelim-plot-particle-type, fig.alt="This bar graph shows the different legnths and kinds of microplastics found within the fish. By overwhelming majority, the most common form of microplastic is single fibers."}
fish_microplastics %>%
  group_by(fish_id) %>%
  drop_na(particle_type) %>%
  summarise(particle_type, count = sum(quantity)) %>%
  ggplot(aes(x = particle_type, y = count)) +
  geom_col()
```
## Reminders:
Ask Laurie about including Leaflet maps in non-HTML outputs
Try to figure out what TWP means in this context

