---
title: "Project proposal"
author: "THE Triangle"
output: github_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(readxl)
library(janitor)
library(broom)
```

## 1. Introduction

Our project will be looking at a 2023 dataset by Leslie Hart of microplastic samples in prey fish caught in Sarasota Bay. The samples were extracted from muscle tissue and GI tracts of the fish, and sorted by size, type, color, and quantity. We are primarily interested in looking at how the geographic information relates to variables like microplastic concentration, color, and body location, and determining if there are any significant correlations.

## 2. Data
Each row represents an observation of microplastic particles in fish tissue. See the data README for codebook and more information.
```{r read-data, message=FALSE}
fish_microplastics <- read_xlsx("../data/Fish_Microplastics_Data_Repository.xlsx", 
                                na = "NA") %>%
  janitor::clean_names() 

```
```{r fix-columns, message=FALSE}
fish_microplastics <- fish_microplastics %>%
  mutate(sieve_size = str_extract(sieve_size, pattern = "[:digit:]+")) %>%
  mutate(sieve_size = case_when(sieve_size == "LG" ~ 777,
          TRUE ~ as.numeric(sieve_size)))
  
```
```{r}
glimpse(fish_microplastics)
```


## 3. Ethics review

## 4. Data analysis plan



